---
title: "Zuur et al. 2009 Mixed Modeling Walkthrough"
subtitle: "Chapter 10"
author: "Michael France Nelson"
date: "6/17/2021"
output: 
  html_notebook:
  toc: true
  toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Packages

```{r packages}
require(here)
require(data.table)
require(DT)

```


## Boars {.tabset}

```{r load_data}
dat_dir = here("book_data", "zuur_et_al_2009")
boars = fread(file.path(dat_dir, "Boar.txt"))
DT::datatable(boars)
```
### Boars plots

```{r boars_exploratory}
{
  par(mfrow = c(1, 2))
  boxplot(LengthCT ~ Tb, data = boars)
  plot(Tb ~ LengthCT, data = boars)
  abline(lm(Tb ~ LengthCT, data = boars))
}

```


### Boars Modeling Goal Plan

1. An assumption on the distribution of the response variable $Y_i$.
* This also defines the mean and variance of $Y_i$.
2. Specification of the systematic part. This is a function of the explanatory variables.
3. The relationship between the mean value of Yi and the systematic part. 
* This is also called the link between the mean and the systematic part.


#### Variables

- The predictor is the head/body length (LengthCT~i~)
- The response is whether or not an individual has TB

#### Distribution of Response

$Y_i$ is $binomial(1, \pi_i)$

- Mean is $1 \times \pi_i$
- Variance is $\pi_i \times (1 - \pi_i)$

#### Predictor Function: Systematic part

Use a linear predictor function $\eta()$.



$\eta(LengthCT_i) = \alpha + \beta \times LengthCT_i$


#### Relationship between expected values and predictors

- Logit link function:
- 'Log of the odds (ratio)'

The odds is the ratio between the probability that it has TB and the probability it does not: $O_i = \frac{\pi_i}{1 - \pi_i}$

Use an exponential function to model the odds:
* In this case we'll use an exponential function for $\eta()$

$O_i = exp( \alpha + \beta \times LengthCT_i) = \frac{\pi_i}{1 - \pi_i}$

And the logarithm:

* $log(O_i) = \alpha + \beta \times LengthCT_i = log(\frac{\pi_i}{1 - \pi_i})$

Shorthand is:

* $log(O_i) = logit(\pi_i) = \alpha + \beta \times LengthCT_i = \eta(X_i)$

Solving for $\pi_i$:

Let $X_i = \alpha + \beta \times LengthCT_i$

* $X_i = log(\frac{\pi_i}{1 - \pi_i})$
* $e^{X_i} = \frac{\pi_i}{1 - \pi_i}$
* $(1 - \pi_i)e^{X_i} = \pi_i$
* $e^{X_i} - \pi_ie^{X_i} = \pi_i$
* $e^{X_i} = \pi_i + \pi_ie^{X_i}$
* $e^{X_i} = \pi_i (1 + e^{X_i})$
* $\pi_i = \frac{e^{X_i}}{1 + e^{X_i}}$
* $e^{X_i} = \frac{\pi_i}{1 - (e^{X_i} - \pi_ie^{X_i})}$

#### Boar GLM

```{r boar_glm_fit}
boar_fit_1 = glm(
  Tb ~ LengthCT, 
  family = binomial,
  data = boars)
summary(boar_fit_1)
```

Plot the model fit

```{r boar_glm_model_fit_plot}
new_dat = data.frame(LengthCT = seq(from = 46.5, to = 165, by = 1))
new_dat$Tb_pred = predict(
  boar_fit_1, newdata = new_dat,
  type = "response")

{
  plot(Tb ~ LengthCT, data = boars)
  lines(Tb_pred ~ LengthCT, data = new_dat)
}
```


# Cod {.tabset}



# GLM for Proportions from counts

## Deer Data



```{r load_deer_data}
deer = fread(file.path(dat_dir, "Tbdeer.txt"))
deer[, DeerNegCervi := (DeerSampledCervi - DeerPosCervi)]
deer[, fFenced := factor(Fenced)]
names(deer)

DT::datatable(deer)

```

### Collinearity Check

Use the authors' function:

```{r}
source(file.path(dat_dir, "HighstatLibV10.R"))

corvif(deer)

```


### Fit a GLM

```{r deer_binomial_glm}
deer_fit_1 = glm(
  cbind(DeerPosCervi, DeerNegCervi) ~
    OpenLand + 
    ScrubLand + 
    QuercusPlants +
    QuercusTrees + 
    ReedDeerIndex + 
    EstateSize + 
    fFenced,
  family = binomial, data = deer)

summary(deer_fit_1)
```


```{r deer_model_collinearity}
require(rms)
require(mctest)


```

