---
title: "Zuur et al. 2009 Mixed Modeling Walkthrough"
subtitle: "Chapter 9: GLM and GAM for Count Data"
author: "Michael France Nelson"
date: "6/17/2021"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```


## Packages and Data

```{r packages_and_data}
require(here)
require(data.table)
require(DT)

dat_dir = here("book_data", "zuur_et_al_2009")
RK = fread(file.path(dat_dir, "RoadKills.txt"))



```


# Generalized Linear Model Structure

Three steps/components:

1. Distribution of the response
2. Specifying the systematic component using the explanatory variables
3. Specifying the link between the model of the means of the response and the systematic part



## Dual Model of the Response Variable

A dual-model of the response:

* Model of mean
* Model of variance

## Systematic Part: Predictor Function $\eta()$



# Poisson GLM Example with log link



## Simulate Data

```{r simulate_counts}

pois_simulator_1 = function(x0, x1, alpha, beta, n = 1)
{
  x = rep(x0:x1, each = n)
  data.frame(
    x = x,
    y = rpois(length(x), lambda = exp(alpha + beta * x)))
}


plot(y ~ x, pois_simulator_1(0, 100, 0.01, 0.03, 2000), pch = 16, col = adjustcolor("steelblue", 0.1))
curve(exp(0.01+ 0.03*x), add = TRUE)

```

Create the data, plot, and fit a model:

```{r plot_sim_data}
dat = pois_simulator_1(0, 200, 0.01, 0.03, 5000)
fit_1 = glm(y ~ x, data = dat, family = poisson)
summary(fit_1)

```




# Road Kill Amphibions


```{r road_kill_model_1}
plot(TOT.N ~ D.PARK, data = RK)
fit_1 = glm(
  TOT.N ~ D.PARK, data = RK, family = poisson)




summary(fit_1)

logLik(fit_1)

```


## Saturated Model

```{r saturated}
RK$saturated = factor(1:nrow(RK))
fit_sat = glm(TOT.N ~ saturated, data = RK, family = poisson)
fit_null = glm(TOT.N ~ 1, data = RK, family = poisson)

summary(fit_sat)
summary(fit_null)
summary(fit_1)

2 * (logLik(fit_sat) - logLik(fit_1))
2 * (logLik(fit_sat) - logLik(fit_null))

```




## Plot model

```{r plot_model_1}
plot(TOT.N ~ D.PARK, data = RK)
x = seq(0, 25000, length.out = 100)
pred_vals = data.frame(
  x,
  predict(
    fit_1,
    newdata = data.frame(
      D.PARK = x),
    se.fit = TRUE,
    type = "response"))
lines(fit ~ x, data = pred_vals)

lines(fit + qnorm(0.025) * se.fit  ~ x, data = pred_vals)
lines(fit - qnorm(0.025) * se.fit  ~ x, data = pred_vals)

```




```{r model_2}
RK$SQ.POLIC <- sqrt(RK$POLIC)
RK$SQ.WATRES <- sqrt(RK$WAT.RES)
RK$SQ.URBAN <- sqrt(RK$URBAN)
RK$SQ.OLIVE <- sqrt(RK$OLIVE)
RK$SQ.LPROAD <- sqrt(RK$L.P.ROAD)
RK$SQ.SHRUB <- sqrt(RK$SHRUB)
RK$SQ.DWATCOUR <- sqrt(RK$D.WAT.COUR)
M2 <- glm(TOT.N ~ OPEN.L + MONT.S + SQ.POLIC +
D.PARK + SQ.SHRUB + SQ.WATRES + L.WAT.C +
SQ.LPROAD + SQ.DWATCOUR, family = poisson,
data = RK)
summary(M2)


```

```{r model_selection_1}


step(M2)
drop1(M2, test = "Chi")
drop1(M2)
```

